Carbonate chemistry at Ny-Ålesund

### 2022-08-09 awipev-CO2.Rmd
fig misc

library("RColorBrewer")
if (!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
if (!require("seacarb")) install.packages("seacarb")
library("seacarb")
if (!require("gridExtra")) install.packages("gridExtra")
library("gridExtra")
if (!require("lubridate")) install.packages("lubridate")
library("lubridate")
if (!require("grid")) install.packages("grid")
library(grid)
if (!require("viridis")) install.packages("viridis")
library("viridis")
if (!require("dygraphs")) install.packages("dygraphs")
library("dygraphs")
if (!require("knitr")) install.packages("knitr")
require("knitr")
if (!require("lmodel2")) install.packages("lmodel2")
library("lmodel2")
if (!require("captioner")) install.packages("captioner")
library("captioner")
if (!require("xts")) install.packages("xts")
library("xts")
if (!require("scales")) install.packages("scales")
library("scales")
if (!require("skimr")) install.packages("skimr")
library("skimr")
if (!require("FNN")) install.packages("FNN")
library("FNN")
if (!require("patchwork")) install.packages("patchwork")
library("patchwork")
if (!require("pander")) install.packages("pander")
library("pander")
if (!require("tsibble")) install.packages("tsibble")
library("tsibble")
if (!require("slider")) install.packages("slider")
library("slider")
if (!require("kableExtra")) install.packages("kableExtra")
library("kableExtra")
if (!require("readxl")) install.packages("readxl")
library("readxl")
if (!require("oce")) install.packages("oce")
library(oce)
if (!require("pangaear")) install.packages("pangaear")
library("pangaear")
if (!require("latex2exp")) install.packages("latex2exp")
library("latex2exp")
if (!require("gtsummary")) install.packages("gtsummary")
library("gtsummary")
if (!require("todor")) install.packages("todor")
library("todor")
# todor::todor_file("awipev-CO2.Rmd")

@@

# Same plot for paper
p_at_s <- d_all_at_discrete %>% 
  drop_na() %>% 
  ggplot(aes(x = sal_fb_mix, y = at)) +
  geom_point(col="blue", size = 0.5) +
  labs(x="Salinity", y=TeX("Total alkalinity ($\\mu mol\\phantom{x}kg^{-1})$"))  +
  stat_smooth(method = "lm", col = "blue", size = 0.5) +
  Mytheme(size_labs = 9)
ggsave(filename = "figures/p_at_s.png", plot = p_at_s, width = 7, height = 7, units = "cm")

@@

## *p*CO~2~ and *f*CO~2~

@@

```{r measured vs calculated pCO2, echo=FALSE, warning=FALSE, message=FALSE, out.width='100%', fig.align="center"}
reg <- ggreg(
  xdata = d_all$pco2_fb_insi_temp,
  ydata = d_all$pco2_calc_fb_insi_temp,
  type = "MA",
  plot = TRUE,
  title = "pCO2 normalized at in situ temperature",
  xlab = latex2exp::TeX("Measured $pCO_{2}\\phantom{x}(\\mu atm)$"),
  ylab = latex2exp::TeX("Calculated $pCO_{2}\\phantom{x}(\\mu atm)$"),
  proplot = TRUE,
  coord_fixed(ratio = 1)
)
print(reg$plot)
pco2_calc_meas <- reg$plot
ggsave(filename = "figures/pco2_calc_meas.png", plot = reg$plot)
```

@@

# ```{r plot measured vs calculated pCO2, echo=FALSE, warning=FALSE, message=FALSE, out.width='100%', fig.align="center"}
# print(reg$plot)
# rm(reg)
# ```

@@

offset_ph <- ph_sf %>% 
  drop_na() %>% 
  ggplot() +
  geom_point(aes(x=datetime, y=diff_spec_sf, color="blue"), size=2) +
  # annotate("text", x = dat_dur$datetime[3], y =-0.16, label = "offset mean", color="red")+
  geom_hline(yintercept=0, size=0.3) +
  geom_hline(yintercept=-0.02, size=0.3, linetype = "dashed") +
  geom_hline(yintercept=0.02, size=0.3, linetype = "dashed") +
  labs(x = NULL, y=expression(paste(pH[T], " difference")), title = NULL) +
  #scale_x_datetime(date_breaks="years", date_minor_breaks="3 months", labels=date_format("%b %Y")) +
  Mytheme() + 
  #theme(panel.grid.major.x = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1)) +
scale_colour_discrete(guide="none")
print(offset_ph)
ggsave(filename = "figures/offset_ph.png", plot = offset_ph, width = 25, height = 20, units = "cm")
```
Offset between calibrated seaFET data and pH spectro. The offset must be between -0.2 and 0.2 pH units. "differences between calibrated sensor pH values and values obtained from discrete, check samples can be expected to be as small as 0.02 pH units, and perhaps less." (McLaughlin et al. 2017).

Initially (before 2022-08-09), there was 2 outiers on 209-07-12 and 2018+09-11. On 2022-08-09, code was addeed above to put a no good qf (3) in order to eliminate those two oiutliers. Now we are fine. The average offset is 'r average_offset'.

@@

```{r pHs comparaison sf spectro, echo=FALSE, message= FALSE, warning=FALSE}
# Spectro / seafet
reg <- ggreg(
  xdata = d_all$ph_s_sf_temp_insi,
  ydata = d_all$pHint_tot_sf,
  type = "MA",
  plot = TRUE,
  title = "pH spectro vs pH seafet (at in situ temperature)",
  xlab = "pH spectro",
  ylab = "pH seafet",
  proplot = TRUE
)
print(reg$plot + xlim(c(8, 8.25)) + ylim(c(8, 8.25)))
```

@@

# identify extremes
d_bp <- d_bp %>%
  dplyr::mutate(
    ext_sal_fb_mix = ifelse(
      sal_fb_mix < dec_sal_fb_mix[1] |
        sal_fb_mix > dec_sal_fb_mix[2],
      sal_fb_mix,
      NA
    ),
    ext_temp_insitu_11m = ifelse(
      temp_insitu_11m < dec_temp_insitu_11m[1] |
        temp_insitu_11m > dec_temp_insitu_11m[2],
      temp_insitu_11m,
      NA
    ),
    ext_pco2_fb = ifelse(pco2_fb < dec_pco2_fb[1] |
                           pco2_fb > dec_pco2_fb[2], pco2_fb, NA),
    ext_pHint_tot_sf = ifelse(
      pHint_tot_sf < dec_pHint_tot_sf[1] |
        pHint_tot_sf > dec_pHint_tot_sf[2],
      pHint_tot_sf,
      NA
    ),
    ext_at_calc = ifelse(at_calc < dec_at_calc[1] |
                           at_calc > dec_at_calc[2], at_calc, NA)
  )
ts_sal_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = sal_fb_mix)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_sal_fb_mix), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="Salinity (< 8 m)", y = "Salinity", x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_temp_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = temp_insitu_11m)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_temp_insitu_11m), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="In situ temperature (11 m)", y = "Temperature (°C)", x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_pco2_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = pco2_fb)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_pco2_fb), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title=expression(paste(pCO[2], " at 11 m and in situ temperature")), y = TeX("pCO$_2$ ($\\mu atm$)"), x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_pH_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = pHint_tot_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_pHint_tot_sf), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="In situ pH below 8 m at in situ temperature", y = expression(paste(pH[T])), x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_at_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = at_calc)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_at_calc), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="Total alkalinity calculated from salinity", y = TeX("$A_T$ ($\\mu mol\\phantom{x}kg^{-1})$"), x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
# Combine plots
g2 <- (ts_sal_bp / ts_temp_bp / ts_pco2_bp / ts_pH_bp / ts_at_bp)
g2 <- g2 + plot_annotation(tag_levels = 'A')
ggsave(filename = "figures/ts_boxplots.png", plot = g2)
#rm(list = ls(pattern = "ts_"), d_bp) #delete big ggplot objects and temp data
#gc() #garbage collection
```
```{r fig ts and bp for paper}
g <- (ts_sal / ts_temp / ts_pCO2  / ts_pH / ts_at) |
  (ts_sal_bp / ts_temp_bp / ts_pco2_bp / ts_pH_bp / ts_at_bp)
g <- g + plot_annotation(tag_levels = 'A')
ggsave(filename = "figures/ts_ts_boxplots.png", plot = g)
rm(list = ls(pattern = "ts_"), d_bp) #delete big ggplot objects and temp data
gc() #garbage collection
```

>>>>

library("RColorBrewer")
if (!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
if (!require("seacarb")) install.packages("seacarb")
library("seacarb")
if (!require("gridExtra")) install.packages("gridExtra")
library("gridExtra")
if (!require("lubridate")) install.packages("lubridate")
library("lubridate")
if (!require("grid")) install.packages("grid")
library(grid)
if (!require("viridis")) install.packages("viridis")
library("viridis")
if (!require("dygraphs")) install.packages("dygraphs")
library("dygraphs")
if (!require("knitr")) install.packages("knitr")
require("knitr")
if (!require("lmodel2")) install.packages("lmodel2")
library("lmodel2")
if (!require("captioner")) install.packages("captioner")
library("captioner")
if (!require("xts")) install.packages("xts")
library("xts")
if (!require("scales")) install.packages("scales")
library("scales")
if (!require("skimr")) install.packages("skimr")
library("skimr")
if (!require("FNN")) install.packages("FNN")
library("FNN")
if (!require("patchwork")) install.packages("patchwork")
library("patchwork")
if (!require("pander")) install.packages("pander")
library("pander")
if (!require("tsibble")) install.packages("tsibble")
library("tsibble")
if (!require("slider")) install.packages("slider")
library("slider")
if (!require("kableExtra")) install.packages("kableExtra")
library("kableExtra")
if (!require("readxl")) install.packages("readxl")
library("readxl")
if (!require("oce")) install.packages("oce")
library(oce)
if (!require("pangaear")) install.packages("pangaear")
library("pangaear")
if (!require("latex2exp")) install.packages("latex2exp")
library("latex2exp")
if (!require("gtsummary")) install.packages("gtsummary")
library("gtsummary")
if (!require("todor")) install.packages("todor")
library("todor")
if (!require("ggpmisc")) install.packages("ggpmisc")
library("ggpmisc")
# todor::todor_file("awipev-CO2.Rmd")

@@

# Same plot for paper
at_s <- d_all_at_discrete %>% 
  drop_na() %>% 
  ggplot(aes(x = sal_fb_mix, y = at)) +
  geom_point(col="blue", size = 0.5) +
  labs(x="Salinity", y=TeX("Total alkalinity ($\\mu mol\\phantom{x}kg^{-1})$"))  +
  stat_smooth(method = "lm", se=FALSE, col = "blue", size = 0.5) +
  Mytheme(size_labs = 9)
ggsave(filename = "figures/at_s.png", plot = at_s, width = 7, height = 7, units = "cm")

@@

## pCO~2~ and fCO~2~

@@

```{r measured vs calculated pCO2, echo=FALSE, warning=FALSE, message=FALSE, out.width='100%', fig.align="center"}
reg <- ggreg(
  xdata = d_all$pco2_fb_insi_temp,
  ydata = d_all$pco2_calc_fb_insi_temp,
  type = "MA",
  plot = TRUE,
  title = "pCO2 normalized at in situ temperature",
  xlab = latex2exp::TeX("Measured $pCO_{2}\\phantom{x}(\\mu atm)$"),
  ylab = latex2exp::TeX("Calculated $pCO_{2}\\phantom{x}(\\mu atm)$"),
  proplot = TRUE,
  coord_fixed(ratio = 1)
)
print(reg$plot)

# for paper
pco2_calc_meas <- d_all %>% 
  dplyr::select(pco2_fb_insi_temp, pco2_calc_fb_insi_temp) %>% 
  drop_na() %>% 
  ggplot(aes(x = pco2_fb_insi_temp, y = pco2_calc_fb_insi_temp)) +
  geom_point(col="blue", size = 0.5) +
  labs(x=latex2exp::TeX("Measured $pCO_{2}\\phantom{x}(\\mu atm)$"),
       y= latex2exp::TeX("Calculated $pCO_{2}\\phantom{x}(\\mu atm)$"))  +
  ggpmisc::stat_ma_line(method = "lmodel2:MA", se = FALSE, size = 1) +
  ggpmisc::stat_ma_eq(aes(label = after_stat(eq.label))) +
#stat_smooth(method = "lm", se=FALSE, col = "blue", size = 0.5) +
  theme_bw() + Mytheme(size_labs = 9)
ggsave(filename = "figures/pco2_calc_meas.png", plot = pco2_calc_meas, width = 25, height = 25, units = "cm")
```

@@

@@

offset_ph <- ph_sf %>% 
  drop_na() %>% 
  ggplot() +
  geom_point(aes(x=datetime, y=diff_spec_sf, color="blue"), size=2) +
  geom_hline(yintercept=0, size=0.3) +
  geom_hline(yintercept=-0.02, size=0.3, linetype = "dashed") +
  geom_hline(yintercept=0.02, size=0.3, linetype = "dashed") +
  labs(x = NULL, y=expression(paste(pH[T], " difference")), title = NULL) +
  Mytheme(size_labs = 12, face_font = "plain") + 
  #theme(panel.grid.major.x = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_colour_discrete(guide="none")
ggsave(filename = "figures/offset_ph.png", plot = offset_ph, width = 10, height = 10, units = "cm")
```
Offset between calibrated seaFET data and pH spectro. The offset must be between -0.2 and 0.2 pH units. "differences between calibrated sensor pH values and values obtained from discrete, check samples can be expected to be as small as 0.02 pH units, and perhaps less." (McLaughlin et al. 2017).

Initially (before 2022-08-09), there was 2 outliers on 2019-07-12 and 2018-09-11. On 2022-08-09, code was added above to put a no good qf (3) in order to eliminate those two outliers. Now we are fine. The average offset is `r round(average_offset, digits = 4)`. It considerably improve the plot "pH spectro vs pH seafet (at in situ temperature)".

@@

### **Spectro vs seafet**
```{r, echo=FALSE, message= FALSE, warning=FALSE}
# Spectro / seafet
reg <- ggreg(
  xdata = d_all$ph_s_sf_temp_insi,
  ydata = d_all$pHint_tot_sf,
  type = "MA",
  plot = TRUE,
  title = "pH spec. vs pH seafet (at in situ temperature)",
  xlab = "Spectrophotometric pH",
  ylab = "SeaFET pH",
  proplot = TRUE
)
pHspec_pHcalc <- reg$plot + xlim(c(8.05, 8.20)) + ylim(c(8.05, 8.20))
p
#print(reg$plot + xlim(c(8, 8.25)) + ylim(c(8, 8.25)))
ggsave(filename = "figures/pHspec_pHcalc.png", plot = pHspec_pHcalc)
```

@@

# identify extremes
d_bp <- d_bp %>%
  dplyr::mutate(
    ext_sal_fb_mix = ifelse(
      sal_fb_mix < dec_sal_fb_mix[1] |
        sal_fb_mix > dec_sal_fb_mix[2],
      sal_fb_mix,
      NA
    ),
    ext_temp_insitu_11m = ifelse(
      temp_insitu_11m < dec_temp_insitu_11m[1] |
        temp_insitu_11m > dec_temp_insitu_11m[2],
      temp_insitu_11m,
      NA
    ),
    ext_pco2_fb = ifelse(pco2_fb < dec_pco2_fb[1] |
                           pco2_fb > dec_pco2_fb[2], pco2_fb, NA),
    ext_pHint_tot_sf = ifelse(
      pHint_tot_sf < dec_pHint_tot_sf[1] |
        pHint_tot_sf > dec_pHint_tot_sf[2],
      pHint_tot_sf,
      NA
    ),
    ext_at_calc = ifelse(at_calc < dec_at_calc[1] |
                           at_calc > dec_at_calc[2], at_calc, NA)
  )
ts_sal_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = sal_fb_mix)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_sal_fb_mix), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="Salinity (< 8 m)", y = "Salinity", x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_temp_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = temp_insitu_11m)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_temp_insitu_11m), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="In situ temperature (11 m)", y = "Temperature (°C)", x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_pco2_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = pco2_fb)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_pco2_fb), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title=expression(paste(pCO[2], " at 11 m and in situ temperature")), y = TeX("pCO$_2$ ($\\mu atm$)"), x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_pH_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = pHint_tot_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_pHint_tot_sf), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="In situ pH below 8 m at in situ temperature", y = expression(paste(pH[T])), x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
ts_at_bp <- d_bp %>%
  ggplot(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = at_calc)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="blue", outlier.shape = NA, na.rm = TRUE) +
  geom_point(aes(x = month(x = datetime, label = TRUE, abbr = TRUE), y = ext_at_calc), col = "lightblue", size = size_point, na.rm = TRUE) +
  labs(title="Total alkalinity calculated from salinity", y = TeX("$A_T$ ($\\mu mol\\phantom{x}kg^{-1})$"), x=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank(),
        aspect.ratio=0.5)
# Combine plots
g2 <- (ts_sal_bp / ts_temp_bp / ts_pco2_bp / ts_pH_bp / ts_at_bp)
g2 <- g2 + plot_annotation(tag_levels = 'A')
ggsave(filename = "figures/ts_boxplots.png", plot = g2)
#rm(list = ls(pattern = "ts_"), d_bp) #delete big ggplot objects and temp data
#gc() #garbage collection
```

###Combined plot for paper

```{r, echo=FALSE, message= FALSE, warning=FALSE}
g <- (at_s / pco2_calc_meas / offset_ph / pHspec_pHcalc)
g <- g + plot_annotation(tag_levels = 'A')
ggsave(filename = "figures/pco2_ph_etc.png", plot = g)
g
```

###Time series and boxplots for paper
```{r }
g <- (ts_sal / ts_temp / ts_pCO2  / ts_pH / ts_at) |
  (ts_sal_bp / ts_temp_bp / ts_pco2_bp / ts_pH_bp / ts_at_bp)
g <- g + plot_annotation(tag_levels = 'A')
ggsave(filename = "figures/ts_ts_boxplots.png", plot = g)
rm(list = ls(pattern = "ts_"), d_bp) #delete big ggplot objects and temp data
gc() #garbage collection
```
